---
title: "EDA: David-M"
format: html
editor: visual
---

## Data Import

```{r}
library(tidyverse)
library(here)
```


```{r}
dating_app_data <- read.csv('../../data/dating_app_behavior_dataset.csv')
```

## Initial Exploration of Data

```{r}
dim(dating_app_data)

head(dating_app_data, 20)
```
```{r}
dating_app_data |>
  group_by(gender) |>
  summarize(gender_count = n())

dating_app_data |>
  group_by(sexual_orientation) |>
  summarize(orientation_count = n())

dating_app_data |>
  group_by(location_type) |>
  summarize(location_count = n())

dating_app_data |>
  group_by(income_bracket) |>
  summarize(income_count = n())

dating_app_data |>
  group_by(education_level) |>
  summarize(education_count = n())
```
::: {.callout-note title="Reflection"}
The dataset has a fairly even distribution across categories, which, while not entirely realistic, allows for more reliable analysis. Since the data is simulated, it’ll be interesting to compare it to actual dating app experiences. 
:::

## Exploring data points individually

```{r}
ggplot(dating_app_data, aes(x = app_usage_time_min)) +
  geom_bar()

ggplot(dating_app_data, aes(x = swipe_right_ratio)) +
  geom_boxplot()
```




## Research Qs

**Research Question:**
What is the most common interest tag?
```{r}
library(stringr)
 
common_tags <- c("music", "hiking", "movies", "travel", "fitness", "food", "books", "dancing", "photography", "art", "tattoos", "yoga", "pets", "DIY", "traveling", "languages", "writing", "painting", "coding")

tag_counts <- sapply(common_tags, function(tag) sum(str_detect(dating_app_data$interest_tags, tag)))

tag_counts_df <- data.frame(tag = common_tags, count = tag_counts)

tag_counts_df <- tag_counts_df[order(-tag_counts_df$count), ]
head(tag_counts_df, 1)

```
```{r}
ggplot(tag_counts_df, aes(x = reorder(tag, count), y = count, fill = tag)) +
  geom_bar(stat = "identity") +
  coord_flip() + 
  labs(title = "Most Common Interest Tags", x = "Interest Tag", y = "Count") +
  theme_minimal() +
  theme(legend.position = "none")
```


**Research Question:**
How does income impact app dependence? 
```{r}
ggplot(dating_app_data, aes(fill=income_bracket,x=app_usage_time_min)) +
         geom_bar(alpha = 0.5)

dating_app_data |>
  group_by(location_type) |>
  mutate(location_mean_usage = mean(app_usage_time_min)) |>
ggplot(aes(x=location_type, y=location_mean_usage)) +
  geom_col()
```

**Research Question:**
Are there income differences in dating app behavior and expectations?
```{r}
# Load required libraries
library(ggplot2)
library(dplyr)

# Ensure income_bracket is an ordered factor
dating_app_data$income_bracket <- factor(dating_app_data$income_bracket, 
                                         levels = c("Very Low", "Low", "Middle", "Upper-Middle", "High"),
                                         ordered = TRUE)

# Create the trivariate plot with facet by match_outcome
ggplot(dating_app_data, aes(x = income_bracket, fill = app_usage_time_label)) +
  geom_bar(position = "dodge") +
  facet_wrap(~match_outcome) +
  labs(title = "Match Outcome by Income and App Usage Time",
       x = "Income Bracket",
       y = "Count",
       fill = "App Usage Time") +
  theme_minimal() +
  theme(legend.position = "top",
        plot.title = element_text(hjust = 0.5))
```

```{r}
# Load required libraries
library(ggplot2)
library(dplyr)

# Ensure income_bracket is an ordered factor
dating_app_data$income_bracket <- factor(dating_app_data$income_bracket, 
                                         levels = c("Very Low", "Low", "Middle", "Upper-Middle", "High"),
                                         ordered = TRUE)

filtered_dates <- dating_app_data %>%
  mutate(match_outcome_clean = tolower(str_trim(match_outcome))) %>%
  filter(!match_outcome_clean %in% c("Catfished", "No Action", "One-sided Like", "Instant Match", "Chat Ignored", "Mutual Match"))

ggplot(filtered_dates, aes(y = income_bracket, fill = app_usage_time_label)) +
  geom_bar(position = "dodge") +
  facet_wrap(~match_outcome) +
  labs(title = "Match Outcomes by Income & App Usage Time",
       x = "Count of People",
       y = "Income Bracket",
       fill = "App Usage Time") +
  theme_minimal() +
  theme(legend.position = "top",
        plot.title = element_text(hjust = 0.5))
```



```{r}
#| echo: false
2 * 2
```

