---
title: "Swiping Through the Spectrum"
number-sections: false
---

```{r}
library(tidyverse)
library(viridisLite)
```

```{r}
dating_app_data <- read.csv("data/dating_app_behavior_dataset.csv")
```

## Introduction

In the modern dating scene, the swipe has become a symbol of possibility, rejection, and everything in between.
But behind every swipe lies a data point---a reflection of who we are, what we seek, and how our identities shape our digital interactions.
This study uses data from a dating app to explore deeper behavioral patterns based on gender, sexual orientation, education4 , income and location.

Why do some users become more "addicted" to dating apps?
How do identity and geography intersect with digital romance?
And what can these patterns tell us about the broader social experience and interaction of dating in the digital age?

## Background

This dataset used in this primary study was a simulation of what a dating app could be.
This data was extracted from Kaggle  We hope our insights will show meaningful patterns which may be used to further develop dating apps.
Our aim is to understand how identity factors relate to app dependency and engagement outcomes, and ultimately explaining broader patterns of human behavior in virtual spaces.

## Analysis

### Gender Identity

We begin with a broad view of how gender identity intersects with app usage levels.

```{r}
dating_app_data |>
  mutate(app_usage_time_label = fct_relevel(app_usage_time_label,
            "Addicted", "Extreme User", "High", "Moderate", "Barely", "Low", "Very Low")) |>
  ggplot(aes(y = gender, fill = app_usage_time_label)) +
  geom_bar(position = "fill") +
  labs(fill = "Usage", y = "Gender") +
  scale_fill_viridis_d(option = "magma") +
  theme_minimal() 
```

The first plot reveals that across all gender identities, the largest usage category is *Extreme User*.
This suggests that intense engagement is a platform-wide phenomenon---dating apps are designed to retain attention across the board (Alter, 2017).

But diving into just the *Addicted* group uncovers something more specific:

```{r, fig.height=3, fig.width=10}
dating_app_data |>
  filter(app_usage_time_label == "Addicted") |>
  group_by(gender) |>
  summarize(count = n()) |>
ggplot(aes(y = gender, x = count, fill = gender)) +
  geom_col() +
  scale_fill_viridis_d(option = "magma") +
  labs(y="Gender", fill="Gender") +
  theme(legend.position = "none") +
  theme_minimal()
```

Here, *Non-binary* (NB) users and those who *Prefer Not to Say* (PNTS) are the most represented.
This is a striking finding and may reflect the fact that NB and PNTS users often find more inclusive spaces online than offline, leading to increased app engagement (Miller, 2015).
Dating apps may act as rare safe zones for connection, identity exploration, and validation.

### Sexual Orientation

Focusing now on NB and PNTS users, we examine how sexual orientation affects app usage time.

```{r}
#| fig-cap: A density visualization depicting the relationship between time spent on app sexual orientation for the gender identities that had the highest rates of "addicted" users as per our analysis with the gender identity factor. 
#| fig-alt: A density visualization depicting the relationship between time spent on app sexual orientation for the gender identities that had the highest rates of "addicted" users as per our analysis with the gender identity factor. 

dating_app_data |>
  filter(app_usage_time_label == "Addicted",
         gender %in% c("Non-binary", "Prefer Not to Say")) |>
  ggplot(aes(y = app_usage_time_min,
             fill = sexual_orientation,
             alpha = 0.2))+
  facet_wrap(~sexual_orientation)+
  geom_density()+
  theme_minimal()+ 
  labs (title = "Distribution of App Usage by Sexual Orientation for NB and PNTS Users", 
        y = "Time Spent on App", 
        x = "Density", 
        fill = "Sexual Orientation")+ 
  guides(alpha = "none")+
  scale_fill_viridis_d()    
```

These plots show that Straight and Queer users within NB/PNTS categories peak in app usage between 55--165 minutes.
The peaks reflect substantial engagement, possibly driven by the limited dating options in offline settings for these orientations (LeFebvre, 2018).

To test this further, we categorize users as Straight vs. Non-Straight:

```{r}
#| fig-cap: A density plot that demonstrates the density of straight vs nonstraight non-binary and "prefer not to say" users as per the minutes they spend on the dating app. 
#| fig-alt: A density plot that demonstrates the density of straight vs nonstraight non-binary and "prefer not to say" users as per the minutes they spend on the dating app. 

dating_app_data |>
  filter( gender %in% c("Non-binary", "Prefer Not to Say")) |>
   mutate(straight = str_detect(sexual_orientation, "Straight")) |>
  ggplot(aes(x = app_usage_time_min,
             fill = straight,
             alpha = 0.5))+
  geom_density()+
  theme_minimal()+ 
  labs (title = "Distribution of App Usage by (Non)Straight NB and PNTS Users", 
        x = "Time Spent on App", 
        y = "Density",
        fill = "User Identifies as Straight")+ 
  guides(alpha = "none")+
  scale_fill_viridis_d()    
```

Interestingly, *Straight* users within NB/PNTS groups show a second peak around 200 minutes, suggesting a higher dependency or reliance on the app.
This could stem from marginalization within both queer and heteronormative spaces, leading to prolonged searching or browsing behavior.

#### What About Binary Genders?

How do these patterns play out among binary gender groups?

```{r}
#How does this relationship change if we look at users that identify as male?
#| fig-cap: A density plot depicting the time usage in minutes of male, straight and male, non-straight users on the dating app.
#| fig-alt: A density plot depicting the time usage in minutes of male, straight and male, non-straight users on the dating app..

dating_app_data |>
  filter(gender == "Male") |>
   mutate(straight = str_detect(sexual_orientation, "Straight")) |>
  ggplot(aes(x = app_usage_time_min,
             fill = straight,
             alpha = 0.5))+
  geom_density()+
  theme_minimal()+ 
  labs (title = "App Usage by (Non)Straight Male Users", 
        x = "Time Spent on App", 
        y = "Density",
        fill = "User Identifies as Straight") + 
  guides(alpha = "none") +
  scale_fill_viridis_d()   
```

For male users, *Straight* individuals spend slightly more time on the app (peak at \~180 mins), possibly due to cultural norms around men being expected to initiate conversations or compete for attention.

```{r}
#| fig-cap: A density plot depicting the time usage in minutes of female, straight vs female, non-straight users on the dating app. 
#| fig-alt: A density plot depicting the time usage in minutes of female, straight vs female, non-straight users on the dating app. 

dating_app_data |>
  filter(gender == "Female") |>
   mutate(straight = str_detect(sexual_orientation, "Straight")) |>
  ggplot(aes(x = app_usage_time_min,
             fill = straight,
             alpha = 0.5))+
  geom_density()+
  theme_minimal()+ 
  labs (title = "App Usage by (Non)Straight Female Users", 
        x = "Time Spent on App", 
        y = "Density",
        fill = "User Identifies as Straight")+ 
  guides(alpha = "none") +
  scale_fill_viridis_d() 
```

Among female users, non-straight individuals spend more time on the app, especially past the 150-minute mark.
This is may be due to the fact that non-straight female users consistently outnumber straight female users beginning at 150 minutes.

### Location

#### Does where you live change how you date?

\[insert plot\]

Users in metro areas spend significantly more time on dating apps than those in suburban or remote areas.
Urban life may facilitate higher engagement due to higher acceptance of app-based dating (Smith & Duggan, 2013).

When comparing average usage time:

\[insert graph here\]

Metro users consistently lead, reinforcing the idea that urban environments foster more digital dating activity whether due to convenience or social norms.
This also could possibly be due to greater social interaction or availability of matches which further supports the idea that people in more densely populated areas tend to engage with dating apps more frequently.

#### Who's Really Matching?

\[insert plot here\]

Our final plot examines mutual matches across location types.
Again, metro users have more variability and higher maximum matches, suggesting greater chances of success in densely populated areas.
The opposite is also true for users in remote areas who appear to have more concentrated and lower values for mutual matches, which suggests a potentially smaller dating pool or fewer interactions.\

### Alternative Explanations

While the data tells a compelling story, alternative interpretations should be considered:

1.  Algorithmic bias may favor urban or frequently active users.

2.  Social isolation in remote areas or among NB/PNTS users may drive higher app engagement.

3.  Psychological dependency like fear of missing out or gamification---may inflate time spent regardless of real connection.

## Conclusion: What the Swipe Says About Us!

Dating apps are more than platforms for meeting people---they're mirrors of modern identity.
Our analysis showed:

-   NB and PNTS users are disproportionately labeled as "Addicted."

-   Urban life correlates with more usage and higher success rates.

-   Patterns differ by gender, sexuality, and geography---but the need for connection unites them all.

To sum it all up, dating apps don't just reveal what we want---they reflect who we are becoming in a digital age.

## References

Huynh, H.
(2018).
Book review: Adam Alter, Irresistible: The rise of addictive technology and the business of keeping us hooked.
Mobile Media & Communication, 7(1), 150--151.
https://doi.org/10.1177/2050157918804400

LeFebvre, L. E.
(2017).
Swiping me off my feet: Explicating relationship initiation on Tinder.
Journal of Social and Personal Relationships, 35(9), 1205--1229.
https://doi.org/10.1177/0265407517706419

Miller, B.
(2015).
"They're the modern-day gay bar": Exploring the uses and gratifications of social networks for men who have sex with men.
Computers in Human Behavior, 51(Part A), 476--482.
https://doi.org/10.1016/j.chb.2015.05.023

Smith, A., & Duggan, M.
(2013, October 21).
Online dating & relationships.
Retrieved from Pew Research Center: Internet, Science & Tech website: https://www.pewresearch.org/internet/2013/10/21/online-dating-relationships/

## Visualizations

<details>

<summary>Click for Visualizations</summary>

```{r}
library(tidyverse)
library(viridisLite)
```

```{r}
dating_app_data <- read.csv("data/dating_app_behavior_dataset.csv")
```

#### Gender

Let's look at user labels.

```{r}
dating_app_data |>
  mutate(app_usage_time_label = fct_relevel(app_usage_time_label,
            "Addicted", "Extreme User", "High", "Moderate", "Barely", "Low", "Very Low")) |>
  ggplot(aes(y = gender, fill = app_usage_time_label)) +
  geom_bar(position = "fill") +
  labs(fill = "Usage", y = "Gender") +
  scale_fill_viridis_d() 
```

The visualization above indicates that the largest usage tag for every gender is "Extreme user".
Now lets look at the "Addicted" category.

```{r, fig.height=5, fig.width=10}
dating_app_data |>
  filter(app_usage_time_label == "Addicted") |>
  group_by(gender) |>
  summarize(count = n()) |>
ggplot(aes(y = gender, x = count, fill = gender)) +
  geom_col() +
  scale_fill_viridis_d() +
  theme(legend.position = "none")
```

So nonbinary users and those who prefer not to say are the largest groups categorized as Addicted.

#### Sexual Orientation

*Note: for this analysis, NB = Non-binary users and PNTS = users that "Prefer Not to Say" for their gender identity.*

```{r}
#| fig-cap: A density visualization depicting the relationship between time spent on app sexual orientation for the gender identities that had the highest rates of "addicted" users as per our analysis with the gender identity factor. 
#| fig-alt: A density visualization depicting the relationship between time spent on app sexual orientation for the gender identities that had the highest rates of "addicted" users as per our analysis with the gender identity factor. 

dating_app_data |>
  filter(app_usage_time_label == "Addicted",
         gender %in% c("Non-binary", "Prefer Not to Say")) |>
  ggplot(aes(x = app_usage_time_min,
             fill = sexual_orientation,
             alpha = 0.2))+
  facet_wrap(~sexual_orientation)+
  geom_density()+
  theme_minimal()+ 
  labs (title = "Distribution of App Usage by Sexual Orientation for NB and PNTS Users", 
        x = "Time Spent on App", 
        y = "Density",
        fill = "Sexual Orientation")+ 
  guides(alpha = "none")+
  scale_fill_viridis_d(option = "magma")    
```

These 8 different density plots demonstrate the difference in time spent on this dating app between users of different sexual orientations.
In this preliminary analysis, we see that there are a few sexual orientations that stand out for having higher densities at higher time spent on the dating app.
Specifically, users that identify as Straight have a density of 0.020 at around \~155 minutes.
Queer users also have a density peak around 165 minutes, reaching a density of \~0.019.

```{r}
#Looking at non-straight and straight users.

#| fig-cap: A density visualization depicting the relationship between time spent on app for straight and non-straight users that held ginder identities of non-binary or "prefer not to say." 
#| fig-alt: A density visualization depicting the relationship between time spent on app for straight and non-straight users that held ginder identities of non-binary or "prefer not to say."

dating_app_data |>
  filter( gender %in% c("Non-binary", "Prefer Not to Say")) |>
   mutate(straight = str_detect(sexual_orientation, "Straight")) |>
  ggplot(aes(x = app_usage_time_min,
             fill = straight,
             alpha = 0.5))+
  geom_density()+
  theme_minimal()+ 
  labs (title = "Distribution of App Usage by (Non)Straight NB and PNTS Users", 
        x = "Time Spent on App", 
        y = "Density",
        fill = "User Identifies as Straight")+ 
  guides(alpha = "none")+
  scale_fill_viridis_d()    
```

Controlling for just users who marked Non-binary or Prefer not to say as their gender identity, this visualization compares users who are straight with those who are not.
For straight users, there is a significant dip around 100 minutes and a higher rate of usage around 200 minutes which could possibly indicate a higher dependency on this dating app for straight NB and PNTS users than others.

```{r}
#How does this relationship change if we look at users that identify as male?
#| fig-cap: A density plot depicting the. 
#| fig-alt: A scatter plot depicting the.

dating_app_data |>
  filter(gender == "Male") |>
   mutate(straight = str_detect(sexual_orientation, "Straight")) |>
  ggplot(aes(x = app_usage_time_min,
             fill = straight,
             alpha = 0.5))+
  geom_density()+
  theme_minimal()+ 
  labs (title = "Distribution of App Usage by (Non)Straight Male Users", 
        x = "Time Spent on App", 
        y = "Density",
        fill = "User Identifies as Straight") + 
  guides(alpha = "none") +
  scale_fill_viridis_d()   
```

Shifting focus slightly, there is a notable difference in time spent on app between non straight and straight male users.
There is a peak in use around 180 minutes for straight males, reaching a density of \~0.004 while non straight males have a density of \~0.0035 at the same amount of time.

```{r}
#| fig-cap: A density plot depicting the. 
#| fig-alt: A scatter plot depicting the.

dating_app_data |>
  filter(gender == "Female") |>
   mutate(straight = str_detect(sexual_orientation, "Straight")) |>
  ggplot(aes(x = app_usage_time_min,
             fill = straight,
             alpha = 0.5))+
  geom_density()+
  theme_minimal()+ 
  labs (title = "App Usage by (Non)Straight Female Users", 
        x = "Time Spent on App", 
        y = "Density",
        fill = "User Identifies as Straight")+ 
  guides(alpha = "none") +
  scale_fill_viridis_d() 
```

Compared to the last visualization, there is less of a distinction between straight and non straight users for female users.
However, there is non-straight female users consistently outnumber straight female users beginning at 150 minutes.
This may possibly indicate that non-straight females spend more time on the app thatn straight females.

#### Location

How does location (i.e. remote area, metro, etc) impact app dependence?

```{r}
dating_app_data |>
  group_by(location_type) |>
  summarise(location_mean_usage = mean(app_usage_time_min, na.rm = TRUE)) |>
  ggplot(aes(x = reorder(location_type, -location_mean_usage), 
             y = location_mean_usage, 
             fill = location_type)) +
  geom_col(show.legend = FALSE) +
  scale_fill_viridis_d(option = "magma") +
  labs(
    title = "Average App Usage Time by Location Type",
    subtitle = "Mean usage time in minutes per location category",
    x = "Location Type",
    y = "Average Usage Time (minutes)"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

In the bar chart above, we can see the average app usage time for different location types.
Metro users consistently show higher average usage compared to those in remote areas, further supporting the idea that people in more densely populated areas tend to engage with dating apps more frequently.

How does location type influence the number of matches or successful interactions users report?

```{r}
library(ggplot2)

ggplot(dating_app_data, aes(x = location_type, y = mutual_matches, fill = location_type)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_viridis_d(option = "magma") +
  labs(
    title = "Distribution of Mutual Matches Across Location Types",
    subtitle = "Boxplot showing variation in mutual match counts by location category",
    x = "Location Type",
    y = "Mutual Matches"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(size = 11, margin = margin(b = 10)),
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 12)
  )

```

The boxplot above displays the distribution of mutual matches by location type.
Metro users tend to have a wider range of mutual matches, with a few high outliers.
This indicates that some users in urban areas may have an exceptional number of successful interactions.
On the other hand, users in remote areas appear to have more concentrated and lower values for mutual matches, which suggests a potentially smaller dating pool or fewer interactions.


