---
title: "Gender"
format: html
execute: 
  echo: false
  warning: false
  error: false
---

```{r}
library(tidyverse)
```

```{r}
dating_app_data <- read.csv('../data/dating_app_behavior_dataset.csv')
```

**First, let's examine the mean usage per gender category**

```{r}
dating_app_data |>
  group_by(gender) |>
  summarize(mean_usage = mean(app_usage_time_min)) |>
ggplot(aes(y=gender, x=mean_usage, fill=gender)) +
  geom_col()
```

**From this, not much difference can be deciphered. Next, we can examine the usage time labels assigned to each.**


```{r}
dating_app_data |>
  mutate(app_usage_time_label = fct_relevel(app_usage_time_label,
            "Addicted", "Extreme User", "High", "Moderate", "Barely", "Low", "Very Low")) |>
  ggplot(aes(y = gender, fill = app_usage_time_label)) +
  geom_bar(position = "fill") +
  labs(fill = "Usage", y = "Gender")
```

**The visualization above indicates that the largest usage tag for every gender is "extreme user". Let's examine only the extreme user category and see the gender differences.**

```{r, fig.height=5, fig.width=10}
dating_app_data |>
  filter(app_usage_time_label == "Addicted") |>
  group_by(gender) |>
  summarize(count = n()) |>
ggplot(aes(y = gender, x = count, fill = gender)) +
  geom_col()
```
**So nonbinary users and those who prefer not to say are the largest groups categorized as Addicted. Let's see how these gorups interact with the app in other ways.**

```{r}
dating_app_data |>
  filter(gender == c("Non-binary", "Prefer Not to Say")) |>
ggplot(aes(x=mutual_matches, fill=gender)) +
  geom_density(alpha = 0.5)
```

```{r}
dating_app_data |>
  filter(gender == c("Non-binary", "Prefer Not to Say")) |>
  ggplot(aes(y=match_outcome, fill = gender)) +
  geom_bar(position = "dodge")
```
**or alternatively**
```{r, fig.width=10}
dating_app_data |>
  filter(gender == c("Non-binary", "Prefer Not to Say")) |>
ggplot(aes(y=match_outcome, fill = gender)) +
  geom_bar() +
  facet_wrap( ~gender)
```
**In most outcomes, nonbinary users had more occurrences for each. However, for date happened, blocked, or mutual match, those who preferred not to say had more of those outcomes.**

```{r}
dating_app_data |>
  filter(gender == c("Male", "Female")) |>
  ggplot(aes(y=match_outcome, fill = gender)) +
  geom_bar(position = "dodge")
```



```{r, fig.height=5, fig.length=10}
# dating_app_data |>
#   mutate(mutual_matches = as.numeric(mutual_matches)) |>
#   mutate(bio_length)
# ggplot(aes(y=bio_length, x=mutual_matches)) +
#   geom_point()
```
